---
title: "analysis"
author: "MYC"
date: "10/15/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---
First loading all the default settings and preliminary programs.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #for data processing
library(here) #to set paths
library(tidymodels)  # for the parsnip package, along with the rest of tidymodels
# Helper packages
library(readr)       # for importing data
library(broom.mixed) # for converting bayesian models to tidy tibbles
```

Path to Processed Data and loading of cleaned data
```{r}
data_location <- here::here("data","processed_data","processeddata.rds")
data<- readRDS(data_location)

glimpse(data)
```

Selection for RunnyNose and previous symptoms
```{r}
analysis<-
  data%>%
  select(BodyTemp, RunnyNose, Nausea, CoughYN,Pharyngitis, Vomit, Diarrhea)
glimpse(analysis)  
```
########
##Model Fitting

Create my linear models using tidymodels

Linear Models
```{r}
lm.model <- linear_reg() %>% set_engine("lm")
```

Linear regression for BodyTemp and RunnyNose
```{r}
lm.fit<- lm.model%>%
  fit(BodyTemp~RunnyNose, analysis)
```


Create clean table
```{r}
tidy(lm.fit)
```
Temp and having a runny nose are affected.

#apply model to all the predictors
```{r}
lm.fit.ALL<-lm.model%>%
  fit(BodyTemp ~., analysis)

tidy(lm.fit.ALL)
```
Bodytemp may effect showing RunnyNose, cough, and pharyngitis 

###
Logistical Models
```{r}
glm.model <- logistic_reg() %>% 
  set_engine("glm")
```

Logistical regression for Nausea and RunnyNose
```{r}
glm.fit<- glm.model%>%
  fit(Nausea~RunnyNose, analysis)

tidy(glm.fit)
```
P>0.05, not significant between Nausea and RunnyNose

#Logistical regression based on Nausea with all predictors
```{r}
glm.fit.ALL<-glm.model%>%
  fit(Nausea~., analysis)
tidy(glm.fit.ALL)
```
P<0.05, significant effects, Vomit and Diarrhea w/ Nausea



