---
title: "EXTRA-AnalysisScriptMod9"
author: "MYC"
date: "10/21/2021"
output: html_document
---
Loading all the default settings and preliminary programs.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #for data processing
library(here) #to set paths
library(tidymodels)      # for the recipes package, along with the rest of tidymodels
library(skimr)           # for variable summaries

```
Path to Processed Data and loading of cleaned data
```{r}
data_location <- here::here("data","processed_data","processeddata.rds")
data<- readRDS(data_location)
```

Selection for RunnyNose and limited symptoms in addition to BodyTemp and Nausea

__Reminder__: Outcome of interest is __Body Temp__; Categorical outcome is __Nausea__; Predictor= __RunnyNose__
```{r}
Mod9Analysis<-
  data%>%
  select(BodyTemp, RunnyNose, Nausea, CoughYN,Pharyngitis, Vomit, Diarrhea)
glimpse(Mod9Analysis)
summary(Mod9Analysis)

```
Changing the dbl data type of BodyTemp into a factor.
97-99 = normal body temperature
anything below or above this range will categorize as abnormal temp
select for this and change BodyTemp to factor
```{r}
Mod9AnalysisFCT<-
  Mod9Analysis%>%
  mutate(
  #convert BodyTemp to a factor
  BodyTemp = ifelse(BodyTemp>= 99, "fever", "normal"),
  BodyTemp = factor(BodyTemp)
  )%>%
  #Exclude missing data
  na.omit()%>%
  # For creating models, it is better to have qualitative columns
  # encoded as factors (instead of character strings)
  mutate_if(is.character, as.factor)

glimpse(Mod9AnalysisFCT)
```
View proportions of the samples with normal temperatures or a fever
```{r}
#making a summary table of proportions
Mod9AnalysisFCT%>%
  count(RunnyNose)%>%
  mutate(Proportions = n/sum(n)) #sample/sum of sample
```
~34% of samples had a fever from this selection of data. The remaining 66% did not present with a fever. 
Most people had normal temperatures in this selection.

```{r}
Mod9AnalysisFCT%>%
  skimr::skim(BodyTemp, RunnyNose)
```
### Data splitting
Need to split into

1. training set
2. testing set

Adapting from the tidymodels example using rsample package to create an object that contains info on how to split and 2 more rsample funtions to create the data frames for training and testing
```{r}
# Fix the random numbers by setting the seed 
# This enables the analysis to be reproducible when random numbers are used 
set.seed(222)
# Put 3/4 of the data into the training set 
data_split <- initial_split(Mod9AnalysisFCT, prop = 3/4)

## Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)

```
##############
# Data has been split
##############
## Create recipies and roles
__roles are optional in this exercise__

### Simple Logistic regression model

Use a recipe to create new predictors and conduct pre-processing required by model

#### Initiate a new recipe
```{r}
Mod9_rec<-
  recipe(BodyTemp~., data = train_data)
#formula to the left of "~" = model outcome
# predictors are to the right of "~"; can be listed of use "." to indicate ALL OTHER variables are predictors
```

View assigned roles whereby Outcome of BodyTemp with RunnyNose as predictor.

#### Add roles to recipe (optional)
```{r}
Mod9_rec<-
  recipe(BodyTemp~., data = train_data)%>%
  update_role(Nausea,CoughYN, Pharyngitis, Vomit, Diarrhea, new_role = "ID")

summary(Mod9_rec)
```

#### Create Features
```{r}
Mod9_rec<-
  recipe(BodyTemp~., data = train_data)%>%
  update_role(Nausea,CoughYN, Pharyngitis, Vomit, Diarrhea, new_role = "ID")%>%
  step_dummy(all_nominal_predictors()) %>% 
  # creates dummy variables as a single factor
    # the all_nominal_predictors to apply the recipe to several variables at once
  step_zv(all_predictors()) 
  #removes columns from data when training set data have a single value

summary(Mod9_rec)
```

#### Fit a model with a recipe
```{r}
#Use logistic regression to model data.
#Build model specification using parsnip
lr_mod<-
  logistic_reg()%>%
  set_engine("glm")
```

Want to use recipe over several steps as training and testing the model.

1. Process the recipe using training set.
2. Apply recipe to training set.
3. Apply recipe to the test set.

To do so use the workflow package from tidymodels
```{r}
Mod9_wflow<-
  workflow()%>%
  add_model(lr_mod)%>%
  add_recipe(Mod9_rec)

Mod9_wflow
```

#### Prepare recipe and train model from resulting predictors
```{r}
Mod9_fit<-
  Mod9_wflow%>%
  fit(data=train_data)
```

Extracting the model or recipe from workflow. Pull fitted model object and see the model coefficients
```{r}
Mod9_fit%>%
  extract_fit_parsnip()%>%
  tidy()
```

## Use a trained workflow to predict unseen test data

Does a Runny Nose correlate with a  fever?
```{r}
predict(Mod9_fit, test_data)
```

```{r}
Mod9_aug<-
  augment(Mod9_fit, test_data)
Mod9_aug
```

```{r}
Mod9_aug%>%
  roc_curve(truth=BodyTemp, .pred_fever)%>%
  autoplot()
```

#### Estimate area under the curve
In general, ROC-AUC =0.5 means the model is no good.
The estimate for this is just barely above this so a RunnyNose isn't a great predictor for a fever if it was at 0.7 it might be useful. so this isn't a great model.
```{r}
Mod9_aug%>%
  roc_auc(truth=BodyTemp, .pred_fever)
```


## Alternative model

Re-do the fitting but with a model that only fits the main predictor to the categorical outcome: _Nausea_.

#### Alterntive recipe
Let's see if having a cough can predict vomiting
```{r}
ALTMod9_rec<-
  recipe(Nausea~., data = train_data)%>%
  update_role(BodyTemp,CoughYN, Pharyngitis, Vomit, Diarrhea, new_role = "ID")

summary(ALTMod9_rec)
```
#### Fit a model with a recipe and workflow

1. Process the recipe using training set.
2. Apply recipe to training set.
3. Apply recipe to the test set.

```{r}
ALTMod9_wflow<-
  workflow()%>%
  add_model(lr_mod)%>%
  add_recipe(ALTMod9_rec)

ALTMod9_wflow
```
#### Prepare recipe and train model from resulting predictors
```{r}
ALTMod9_fit<-
  ALTMod9_wflow%>%
  fit(data=train_data)
```

Extracting the model or recipe from workflow. Pull fitted model object and see the model coefficients
```{r}
ALTMod9_fit%>%
  extract_fit_parsnip()%>%
  tidy()
```
## Use a trained workflow to predict unseen test data

If runny noses result in nausea

Predict returns predicted class Yes or No
```{r}
predict(ALTMod9_fit, test_data)
```
Using augment with the model plus test data to save prediction
```{r}
ALTMod9_aug<-
  augment(ALTMod9_fit, test_data)
ALTMod9_aug
```
Generate and ROC curve.
Uses the probability of Nausea being present with a Runny nose
```{r}
ALTMod9_aug%>%
  roc_curve(truth=Nausea, .pred_Yes)%>%
  autoplot()
```

```{r}
ALTMod9_aug%>%
  roc_auc(truth=Nausea, .pred_Yes)
```


###############################################################
